% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\usepackage{fvextra}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
\DefineVerbatimEnvironment{OutputCode}{Verbatim}{breaklines,commandchars=\\\{\}}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Referee report},
  pdfauthor={Hyoungchul Kim},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{Referee report}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Fragmented markets and the proliferation of small firms:
Evidence from mom-and-pop shops in Mexico (Ramos-Menichelli and
Sverdlin-Lisker)}
\author{Hyoungchul Kim}
\date{2025-04-18}

\begin{document}
\maketitle


\subsection{Summary}\label{summary}

This paper studies the proliferation of small firms in developing
countries by focusing on Mexico's traditional retail sector. The authors
develop a spatial model in which high consumer transport costs lead to
fragmented markets, reducing effective market size and encouraging entry
by small, low-quality firms, particularly when entry costs are low. The
model generates clear comparative statics and motivates an empirical
analysis that uses the 2017 deregulation of Mexico's gasoline market as
a source of exogenous variation in transport costs.

Combining this policy shock with a novel panel of firm-level data
covering the universe of mom-and-pop shops in Mexico, the authors show
that areas facing higher gas price increases experienced a rise in the
number of stores, a decline in average store size, and a fall in average
store quality. The paper then uses the estimated model to evaluate the
welfare consequences of a licensing policy that increases entry costs,
and finds modest welfare gains from reducing firm entry.

The paper makes several important contributions:

\begin{itemize}
\item
  It presents a clear and tractable spatial model that incorporates firm
  heterogeneity and consumer mobility, offering a new perspective on
  market fragmentation in developing economies.
\item
  The empirical strategy leverages a high-frequency, spatially detailed
  dataset that captures informal firms typically missed in
  administrative data, and exploits a credible source of exogenous
  variation.
\item
  The paper adds to a growing literature emphasizing the role of
  demand-side frictions in shaping firm dynamics and market structure.
\item
  Finally, it provides policy-relevant insights by evaluating the
  trade-offs of a licensing policy that reduces excessive entry.
\end{itemize}

\section{Comments}\label{comments}

\subsection{Major comments}\label{major-comments}

\begin{itemize}
\item
  Even though the author talks about this matter, it would still be nice
  if the author could elaborate bit more on the importance of
  transportation cost on market segmentation in developing countries
  compared to developed countries. My prior on people's consumption
  behavior is that people usually shop necessity goods near their
  residence (walkable distance). In that sense, I am not quite sure how
  transportation cost would significantly segment the market. I am
  thinking this is mostly due to the differences in the economic
  environment in developing cities. Thus it would be nice for the paper
  to talk more about this.
\item
  As author mentioned that consumers and firms might have anticipated
  the energy reform policy, it might be nice to vary the treatment
  timing to check that the results do not change much. For example, it
  might be good to set the timing to previous dates if the news about
  the policy arrived before the actual implementation of the policy. If
  the author believe that there would not have been such anticipation
  effect, anecdotal evidence would also suffice.
\item
  The model treats supermarket behavior and location as fixed. While
  this may be reasonable over a short horizon, some discussion of
  longer-run dynamics or potential general equilibrium effects (e.g.,
  how upstream suppliers or supermarkets might respond) would be useful.
\item
  The authors argue that the findings are relevant to other developing
  countries. A short discussion of where this framework is most likely
  to apply (e.g., settings with similarly low entry costs and poor
  transport infrastructure) would strengthen the broader relevance of
  the results.
\end{itemize}

\subsection{Minor comments}\label{minor-comments}

\begin{itemize}
\item
  It might be better to use poisson regression when analyzing impact of
  transportation cost on firm opening and closings because arbitrary log
  transformation of inverst hyperbolic sine transformation are not
  recommended in the current literature.
\item
  For figure 5, it might be better to plot it in histogram which would
  make it more consistent with figure 3 which also plot by income
  decile.
\item
  For figure 6, it might be better to later use gradient color scheme
  rather than different colors (blue, yellow, red) to distinguish them
  in case of black and white format.
\item
  For figure 14 and 16, it might be better to zoom in on the values and
  the confidence intervals because currently it is taking too small
  space in the y range, making it visually harder to interpret.
\item
  The literature review is broad and well-integrated, but it could more
  clearly position this paper in contrast to closely related work (e.g.,
  Jensen and Miller (2018); You 2021).
\end{itemize}




\end{document}
