---
title: "CSDID"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "2023-03-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\section{Second Part}
\textbf{I also ask you to replicate the simulation on Sant’anna’s website which shows that TWFE event
studies can be quite biased in the case of staggered designs. The simulation and results are in a blog here: https://psantanna.com/posts/twfe# . He has also put up the R code to replicate these
results. For the problem set we will start with the runs where the data includes a never treated group and when treatment effects are heterogeneous across groups and over time (this is the last scenario he discusses).}

\subsection*{Question 1}

\textbf{Simulate the data and plot the raw trends in values by group over time. Check that it looks like his figure.}

```{r}
library(tidyverse)
library(lfe)
library(fastDummies)
library(ggthemes)
library(ggplot2)
library(did)
library(tibble)
library(rlang)
#install.packages("devtools")
library(conflicted)
#library(dplyr)
theme_set(theme_clean() + theme(plot.background = element_blank()))
#----------------------------------------------------------------------------
iseed  = 20201221
nrep <- 100  
true_mu <- 1
set.seed(iseed)
#pdflatex()
```

```{r warning=FALSE, include=FALSE}
options(tinytex.verbose = TRUE)
```

```{r}
## Generate data - treated cohorts consist of 250 obs each, with the treatment effect still = true_mu on average
make_data3 <- function(nobs = 1000, 
                      nstates = 40) {
  
  # unit fixed effects (unobservd heterogeneity)
  unit <- tibble(
    unit = 1:nobs,
    # generate state
    state = sample(1:nstates, nobs, replace = TRUE),
    unit_fe = rnorm(nobs, state/5, 1),
    # generate instantaneous treatment effect
    #mu = rnorm(nobs, true_mu, 0.2)
    mu = true_mu
  )
  
  # year fixed effects (first part)
  year <- tibble(
    year = 1980:2010,
    year_fe = rnorm(length(year), 0, 1)
  )
  
  # Put the states into treatment groups
  treat_taus <- tibble(
    # sample the states randomly
    state = sample(1:nstates, nstates, replace = FALSE),
    # place the randomly sampled states into 1\{t \ge g \}G_g
    cohort_year = sort(rep(c(1986, 1992, 1998, 2004), 10))
  )
  
  # make main dataset
  # full interaction of unit X year 
  expand_grid(unit = 1:nobs, year = 1980:2010) %>% 
    left_join(., unit) %>% 
    left_join(., year) %>% 
    left_join(., treat_taus) %>% 
    # make error term and get treatment indicators and treatment effects
    # Also get cohort specific trends (modify time FE)
    mutate(error = rnorm(nobs*31, 0, 1),
           treat = ifelse((year >= cohort_year)* (cohort_year != 2004), 1, 0),
           mu = ifelse(cohort_year==1992, 2, ifelse(cohort_year==1998, 1, 3)),
           tau = ifelse(treat == 1, mu, 0),
           year_fe = year_fe + 0.1*(year - cohort_year)
    ) %>% 
    # calculate cumulative treatment effects
    group_by(unit) %>% 
    mutate(tau_cum = cumsum(tau)) %>% 
    ungroup() %>% 
    # calculate the dep variable
    mutate(dep_var = (2010 - cohort_year) + unit_fe + year_fe + tau_cum + error) %>%
    # Relabel 2004 cohort as never-treated
    mutate(cohort_year = ifelse(cohort_year == 2004, Inf, cohort_year))
  
}
#----------------------------------------------------------------------------
# make data
data <- make_data3()

# plot
plot3 <- data %>% 
  ggplot(aes(x = year, y = dep_var, group = unit)) + 
  geom_line(alpha = 1/8, color = "grey") + 
  geom_line(data = data %>% 
              group_by(cohort_year, year) %>% 
              summarize(dep_var = mean(dep_var)),
            aes(x = year, y = dep_var, group = factor(cohort_year),
                color = factor(cohort_year)),
            linewidth = 2) + 
  labs(x = "", y = "Value",  color = "Treatment group   ") + 
  geom_vline(xintercept = 1986, color = '#E41A1C', size = 2) + 
  geom_vline(xintercept = 1992, color = '#377EB8', size = 2) + 
  geom_vline(xintercept = 1998, color = '#4DAF4A', size = 2) + 
  #geom_vline(xintercept = 2004, color = '#984EA3', size = 2) + 
  scale_color_brewer(palette = 'Set1') + 
  theme(legend.position = 'bottom',
        #legend.title = element_blank(), 
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12)) +
  scale_color_manual(labels = c("1986", "1992", "1998", "Never-treated"),
                     values = c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3")) +
  ggtitle("One draw of the DGP with heterogeneous treatment effect dynamics across cohorts \n and with a never-treated group")+
  theme(plot.title = element_text(hjust = 0.5, size=12))

plot3

```

\subsection*{Question 2}
\textbf{Estimate the TWFE event study with all leads and lags (i.e., no binning). Does it approximately recover the true treatment effects in event time?}
```{r}
# function to run ES DID
# variables we will use
keepvars <- c("`rel_year_-5`",  "`rel_year_-4`",  "`rel_year_-3`",  "`rel_year_-2`",
              "rel_year_0", "rel_year_1", "rel_year_2", "rel_year_3", "rel_year_4", "rel_year_5")

run_ES_DiD_never_het <- function(...) {
  
  # resimulate the data
  data <- make_data3()
  # make dummy columns
  data <- data %>% 
    # make dummies
    mutate(rel_year = year - cohort_year) %>% 
    mutate(rel_year = ifelse(rel_year == -Inf, NA, rel_year))%>% 
    dummy_cols(select_columns = "rel_year") %>% 
    mutate(across(starts_with("rel_year_"), ~replace_na(., 0))) %>% 
    # generate pre and post dummies
    mutate(Pre = ifelse((rel_year < -5) * (!is.na(rel_year)), 1, 0),
           Post = ifelse((rel_year > 5) * (!is.na(rel_year)), 1, 0)) %>%
    mutate(Pre = ifelse(is.na(Pre), 0, Pre),
           Post = ifelse(is.na(Post), 0, Post))
  
  # estimate the model
  #mod <- lfe::felm(dep_var ~ Pre + `rel_year_-5` + `rel_year_-4` + `rel_year_-3` + `rel_year_-2` + 
  #              `rel_year_0` + `rel_year_1` + `rel_year_2` + `rel_year_3` + `rel_year_4` + 
  #              `rel_year_5` + Post | unit + year | 0 | state, data = data, exactDOF = TRUE)

  mod <- lfe::felm(dep_var ~  `rel_year_-18`+ `rel_year_-17`+ `rel_year_-16`+ `rel_year_-15`+ `rel_year_-14`+ `rel_year_-13` + `rel_year_-12` + `rel_year_-11` + `rel_year_-10` + `rel_year_-9` + `rel_year_-8` + `rel_year_-7`+ `rel_year_-6` + `rel_year_-5` + `rel_year_-4`+ `rel_year_-3` + `rel_year_-2` + `rel_year_0` + `rel_year_1` + `rel_year_2` + `rel_year_3` + `rel_year_4` + `rel_year_5`+ `rel_year_6`+ `rel_year_7`+ `rel_year_8`+ `rel_year_9` + `rel_year_10` + `rel_year_11`+ `rel_year_12`| unit + year | 0 | state, data = data, exactDOF = TRUE)
  #`rel_year_-19`+ ` rel_year_-20`+ `rel_year_-21`+ `rel_year_-22`+ `rel_year_-23`+ `rel_year_-24`
  
  # grab the obs we need
# grab the obs we need
  mod2 <- tibble(
    estimate = mod$coefficients,
    term1 = rownames(mod$coefficients)
    )
 es <-
   mod2 %>% 
    #filter(term1 %in% keepvars) %>% 
    mutate(t = c(-18:-2, 0:12)) %>% 
    select(t, estimate)
 #es <-
#   mod2 %>% 
    #filter(term1 %in% keepvars) %>% 
 #   mutate(t = c(-5:-2, 0:5)) %>% 
  #  select(t, estimate)
 es
}

data_classical_never_het <- map_dfr(1:nrep, run_ES_DiD_never_het)

ES_plot_classical_never_het <- data_classical_never_het %>% 
  group_by(t) %>% 
  summarize(avg = mean(estimate),
            sd = sd(estimate),
            lower.ci = avg - 1.96*sd,
            upper.ci = avg + 1.96*sd) %>% 
  bind_rows(tibble(t = -1, avg = 0, sd = 0, lower.ci = 0, upper.ci = 0)) %>% 
  mutate(true_tau = ifelse(t >= 0, (t + 1)* 2, 0)) %>% 
  ggplot(aes(x = t, y = avg)) + 
  #geom_linerange(aes(ymin = lower.ci, ymax = upper.ci), color = 'darkgrey', size = 2) + 
  geom_ribbon(aes(ymin = lower.ci, ymax = upper.ci), color = "lightgrey", alpha = 0.2) +
  geom_point(color = 'blue', size = 3) + 
   geom_line(aes(color = 'Estimated Effect'), size = 1) + 
   geom_line(aes(x = t, y = true_tau, color = 'True Effect'), linetype = "dashed", size = 2) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  scale_x_continuous(breaks = -19:12) + 
  labs(x = "Relative Time", y = "Estimate") + 
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12))+
  ggtitle("TWFE event-study regression NOT with binned end-points")+
  scale_color_manual(values = colors) + 
  theme(plot.title = element_text(hjust = 0.5, size=12),
        legend.position = "bottom", 
        legend.title = element_blank())

ES_plot_classical_never_het
```


```{r}
# function to run ES DID
run_ES_DiD_sat_never_het <- function(...) {
  
  # resimulate the data
  data <- make_data3()
  
  # make dummy columns
  data <- data %>% 
    # make relative year indicator
    mutate(rel_year = year - cohort_year)
  
  # get the minimum relative year - we need this to reindex
  min_year <- min(data$rel_year * (data$rel_year != -Inf), na.rm = T)
  
  # reindex the relative years
  data <- data %>% 
    mutate(rel_year2 = rel_year) %>% 
    mutate(rel_year = rel_year - min_year) %>% 
    dummy_cols(select_columns = "rel_year") %>% 
    select(-("rel_year_-Inf"))
    
  
  # make regression formula 
  indics <- paste("rel_year", (1:max(data$rel_year))[-(-1 - min_year)], sep = "_", collapse = " + ")
  keepvars <- paste("rel_year", c(-5:-2, 0:5) - min_year, sep = "_")  
  formula <- as.formula(paste("dep_var ~", indics, "| unit + year | 0 | state"))
  
  # run mod
  mod <- felm(formula, data = data, exactDOF = TRUE)
  
  # grab the obs we need
# grab the obs we need
  mod2 <- tibble(
    estimate = mod$coefficients,
    term1 = rownames(mod$coefficients)
    )
  
 es <-
   mod2 %>% 
    filter(term1 %in% keepvars) %>% 
    mutate(t = c(-5:-2, 0:5)) %>% 
    select(t, estimate)
 es
}

data_sat_never_het <- map_dfr(1:nrep, run_ES_DiD_sat_never_het)

ES_plot_sat_never_het <- data_sat_never_het %>% 
  group_by(t) %>% 
  summarize(avg = mean(estimate),
            sd = sd(estimate),
            lower.ci = avg - 1.96*sd,
            upper.ci = avg + 1.96*sd) %>% 
  bind_rows(tibble(t = -1, avg = 0, sd = 0, lower.ci = 0, upper.ci = 0)) %>% 
  mutate(true_tau = ifelse(t >= 0, (t + 1)* 2, 0)) %>% 
  ggplot(aes(x = t, y = avg)) + 
  #geom_linerange(aes(ymin = lower.ci, ymax = upper.ci), color = 'darkgrey', size = 2) + 
  geom_ribbon(aes(ymin = lower.ci, ymax = upper.ci), color = "lightgrey", alpha = 0.2) +
  geom_point(color = 'blue', size = 3) + 
   geom_line(aes(color = 'Estimated Effect'), size = 1) + 
   geom_line(aes(x = t, y = true_tau, color = 'True Effect'), linetype = "dashed", size = 2) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  scale_x_continuous(breaks = -5:5) + 
  labs(x = "Relative Time", y = "Estimate") + 
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12))+
  ggtitle("TWFE event-study regression with 'all' leads and lags")+
  scale_color_manual(values = colors) + 
  theme(plot.title = element_text(hjust = 0.5, size=12),
        legend.position = "bottom", 
        legend.title = element_blank())

ES_plot_sat_never_het

```

\subsection*{Question 3}
\textbf{Now use the Callaway Sant’anna estimator to produce the event study. Does this match the true effect?}
```{r}
run_CS_never_het <- function(...) {
  
  # resimulate the data
  data <- make_data3()
  data$cohort_year[data$cohort_year==Inf] <- 0
  
  mod <- did::att_gt(yname = "dep_var", 
                     tname = "year",
                     idname = "unit",
                     gname = "cohort_year",
                     control_group= "nevertreated",
                     bstrap = FALSE,
                     data = data,
                     print_details = FALSE)
  event_std <- did::aggte(mod, type = "dynamic")
  
  att.egt <- event_std$att.egt
  names(att.egt) <- event_std$egt
  
  # grab the obs we need
  broom::tidy(att.egt) %>% 
    filter(names %in% -5:5) %>% 
    mutate(t = -5:5, estimate = x) %>% 
    select(t, estimate)
}

data_CS_never_het <- map_dfr(1:nrep, run_CS_never_het)

ES_plot_CS_never_het <- data_CS_never_het %>% 
  group_by(t) %>% 
  summarize(avg = mean(estimate),
            sd = sd(estimate),
            lower.ci = avg - 1.96*sd,
            upper.ci = avg + 1.96*sd) %>% 
  mutate(true_tau = ifelse(t >= 0, (t + 1)* 2, 0)) %>% 
  ggplot(aes(x = t, y = avg)) + 
  #geom_linerange(aes(ymin = lower.ci, ymax = upper.ci), color = 'darkgrey', size = 2) + 
  geom_ribbon(aes(ymin = lower.ci, ymax = upper.ci), color = "lightgrey", alpha = 0.2) +
  geom_point(color = 'blue', size = 3) + 
   geom_line(aes(color = 'Estimated Effect'), size = 1) + 
   geom_line(aes(x = t, y = true_tau, color = 'True Effect'), linetype = "dashed", size = 2) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  scale_x_continuous(breaks = -5:5) + 
  labs(x = "Relative Time", y = "Estimate") + 
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12))+
  ggtitle("Event-study-parameters estimated using Callaway and Sant'Anna (2020)\nComparison group: Never-treated units")+
  scale_color_manual(values = colors) + 
  theme(plot.title = element_text(hjust = 0.5, size=12),
        legend.position = "bottom", 
        legend.title = element_blank())

ES_plot_CS_never_het

```


\subsection*{Question 4}
\textbf{Now, let’s change the data generating process so that the treatment effects are constant over time and across all groups. So, the outcome jumps on treatment but stays stable over time instead of increasing. Re-estimate the TWFE estimator on this new constant treatment effects data. This time does the TWFE recover the true effects? }

```{r}
make_data <- function(nobs = 1000, 
                      nstates = 40) {
  
  # unit fixed effects (unobservd heterogeneity)
  unit <- tibble(
    unit = 1:nobs,
    # generate state
    state = sample(1:nstates, nobs, replace = TRUE),
    unit_fe = rnorm(nobs, state/5, 1),
    # generate instantaneous treatment effect
    #mu = rnorm(nobs, true_mu, 0.2)
    mu = true_mu
  )
  
  # year fixed effects (first part)
  year <- tibble(
    year = 1980:2010,
    year_fe = rnorm(length(year), 0, 1)
  )
  
  # Put the states into treatment groups
  treat_taus <- tibble(
    # sample the states randomly
    state = sample(1:nstates, nstates, replace = FALSE),
    # place the randomly sampled states into four treatment groups G_g
    cohort_year = sort(rep(c(1986, 1992, 1998, 2004), 10))
  )
  
  # make main dataset
  # full interaction of unit X year 
  expand_grid(unit = 1:nobs, year = 1980:2010) %>% 
    left_join(., unit) %>% 
    left_join(., year) %>% 
    left_join(., treat_taus) %>% 
    # make error term and get treatment indicators and treatment effects
    # Also get cohort specific trends (modify time FE)
    mutate(error = rnorm(nobs*31, 0, 1),
           treat = ifelse(year >= cohort_year, 1, 0),
           tau = ifelse(treat == 1, mu, 0),
           year_fe = year_fe + 0.1*(year - cohort_year)
    ) %>% 
  
    # calculate the dep variable
    mutate(dep_var = (2010 - cohort_year) + unit_fe + year_fe + tau + error)
  
}

data_classical <- map_dfr(1:nrep, run_ES_DiD) 

colors <- c("True Effect" = "red", "Estimated Effect" = "blue")

ES_plot_classical <- data_classical %>% 
  group_by(t) %>% 
  summarize(avg = mean(estimate),
            sd = sd(estimate),
            lower.ci = avg - 1.96*sd,
            upper.ci = avg + 1.96*sd) %>% 
  bind_rows(tibble(t = -1, avg = 0, sd = 0, lower.ci = 0, upper.ci = 0)) %>% 
  mutate(true_tau = ifelse(t >= 0, true_mu, 0)) %>% 
  ggplot(aes(x = t, y = avg)) + 
  #geom_linerange(aes(ymin = lower.ci, ymax = upper.ci), color = 'darkgrey', size = 2) + 
  geom_ribbon(aes(ymin = lower.ci, ymax = upper.ci), color = "lightgrey", alpha = 0.2) +
  geom_point(color = 'blue', size = 3) + 
  geom_line(aes(color = 'Estimated Effect'), size = 1) + 
  geom_line(aes(x = t, y = true_tau, color = 'True Effect'), linetype = "dashed", size = 2) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  scale_x_continuous(breaks = -5:5) + 
  labs(x = "Relative Time", y = "Estimate") + 
  theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12)) +
  ggtitle("TWFE event-study regression with binned end-points")+
  scale_color_manual(values = colors) + 
  theme(plot.title = element_text(hjust = 0.5, size=12),
        legend.position = "bottom", 
        legend.title = element_blank())

ES_plot_classical 
```



